---
title: "ECON 490 Data Analysis"
author: 'Shijia Huang'
date: "`r Sys.Date()`"
output: pdf_document
---

```{r set-up, include=FALSE}
library(tidyverse)
library(lubridate)
library(twitteR)
library(tm)
library(wordcloud)
library(syuzhet)
library(zoo)
library(maps)
library(ggthemes)
library(dplyr)
library(ggpubr)
library(AER)
library(plm)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = TRUE)
```








# Part 1: Data Preprocessing and Cleaning (long running time!!!)

```{r}
# Load raw datasets
tweets <- read.csv("covid_vaccine_v36.csv", header = TRUE)
vaccine <- read.csv("covid_vaccine_admistrated.csv", header = TRUE)
cases <- read.csv("covid_cases.csv", header = TRUE)
cities <- read.csv("uscities.csv", header = TRUE)
```


## Tweets data cleaning

```{r}
# Cleaning Date: convert chr date and time to date object
tweets <- tweets[!tweets$date == "", ]
tweets$date <- as.Date(parse_date_time(tweets$date, orders = c('dmY HM', 'Ymd HMS'), tz = "UTC"))
tweets <-  tweets[!is.na(tweets$date), ]

# Cleaning and match user_location
# remove empty user_location rows
tweets <- tweets[!tweets$user_location == "", ]
# filter and cleaning location only in US cities
# Find the 10 most frequent locations
tail(names(sort(table(tweets$user_location))), 10)

# Roughly remove rows where user_location not in US
rmv <- c("India", "United Kingdom", "England", "London", "UK", "Canada", "Australia")
for(i in 1:length(rmv)) {
  tweets <- filter(tweets, str_detect(user_location, regex(rmv[i], ignore_case = T), 
                                      negate = TRUE))
}

# get rid of all non-ASCII characters
tweets$user_location <- sapply(tweets$user_location,
                               function(row) iconv(row, "latin1", "ASCII", sub=""))
tweets <- tweets[!tweets$user_location == "", ]
```

### 1. Strict match location by the format City, State

```{r}
tweets <- add_column(tweets, user_state = rep(NA), .after = "user_location")
format1 <- str_c(cities$city_ascii, cities$state_id, sep = ", ", collapse = NULL)
cities <- add_column(cities, format1 = format1, .after = "state_name")

for (i in 1:length(tweets$user_location)) {
  lca <- tweets$user_location[i]
  j <- match(lca, cities$format1)
  state <- cities$state_id[j]
  tweets$user_state[i] <- state
}


tweets_left <- tweets[is.na(tweets$user_state),]
tweets <- tweets[!is.na(tweets$user_state),]

# see what we get
ggplot(tweets, aes(user_state)) + geom_bar() + coord_flip()
```

### 2. Strict match location by the format City,State

```{r}
format2 <- str_c(cities$city_ascii, cities$state_id, sep = ",", collapse = NULL)
cities <- add_column(cities, format2 = format2, .after = "format1")

for (i in 1:length(tweets_left$user_location)) {
  lca <- tweets_left$user_location[i]
  j <- match(lca, cities$format2)
  state <- cities$state_id[j]
  tweets_left$user_state[i] <- state
}

tweets <- rbind(tweets, tweets_left[!is.na(tweets_left$user_state),])
tweets_left <- tweets_left[is.na(tweets_left$user_state),]

# see what we get now
ggplot(tweets, aes(user_state)) + geom_bar() + coord_flip()
```

### 3. Strict match location by the State Name

```{r}
for (i in 1:length(tweets_left$user_location)) {
  lca <- tweets_left$user_location[i]
  j <- match(lca, cities$state_name)
  state <- cities$state_id[j]
  tweets_left$user_state[i] <- state
}

tweets <- rbind(tweets, tweets_left[!is.na(tweets_left$user_state),])
tweets_left <- tweets_left[is.na(tweets_left$user_state),]

# see what we get now
ggplot(tweets, aes(user_state)) + geom_bar() + coord_flip()
```

### 4. Strict match location by the format State, USA

```{r}
format3 <- str_c(cities$state_name, "USA", sep = ", ", collapse = NULL)
states <- unique(cbind.data.frame(cities$state_id, cities$state_name, format3))
colnames(states) <- c("state_id", "state_name", "format3")

for (i in 1:length(tweets_left$user_location)) {
  lca <- tweets_left$user_location[i]
  j <- match(lca, states$format3)
  state <- states$state_id[j]
  tweets_left$user_state[i] <- state
}

tweets <- rbind(tweets, tweets_left[!is.na(tweets_left$user_state),])
tweets_left <- tweets_left[is.na(tweets_left$user_state),]

# see what we get now and save the filtered file
ggplot(tweets, aes(user_state)) + geom_bar() + coord_flip()

#write.csv(tweets, "tweets_cleaned.csv", row.names=FALSE)
```

### Cleaning tweets text

```{r}
# reload cleaned file
tweets <- read.csv("tweets_cleaned.csv", header = TRUE)
tweets$date <- as.Date(tweets$date)

# Convert all text to lower case
tweets$text <- iconv(tweets$text, "latin1", "ASCII", "")
tweets$text <- tolower(tweets$text)

# Replace @UserName
tweets$text <- gsub("@\\w+", "", tweets$text)

# Remove punctuation
tweets$text <- gsub("[[:punct:]]", "", tweets$text)

# Remove links
tweets$text <- gsub("http\\w+", "", tweets$text)

# Replace tabs
tweets$text <- gsub("[ |\t]{2,}", " ", tweets$text)

# Replace blank space
tweets$text <- gsub("\\s+"," ", tweets$text)

# Remove blank spaces at the beginning
tweets$text <- gsub("^ ", "", tweets$text)

# Remove blank spaces at the end
tweets$text <- gsub(" $", "", tweets$text)

# Drop NA rows
tweets <- tweets[!is.na(tweets$text), ]
tweets <- tweets[!tweets$text == "", ]
```


## Vaccine data cleaning

```{r}
# Convert chr date and time to datetime
vaccine$Date <- as.Date(parse_date_time(vaccine$Date, 'mdY', tz = "UTC"))
```


## Cases data cleaning

```{r}
# Convert chr date and time to datetime
cases$submission_date <- as.Date(parse_date_time(cases$submission_date, 'mdY', tz = "UTC"))
```


## Combine datasets 

```{r}
tweets <- subset(tweets, select = c(user_name, user_location, user_state, date, 
                                    text, hashtags))
cases <- subset(cases, select = c(submission_date, state, tot_cases, new_case, tot_death, 
                                  new_death))
vaccine <- subset(vaccine, select = c(Date, Location, Admin_Per_100K))

# merge and save the merged data
tweets_cases <- merge(tweets, cases, by.x = c("date", "user_state"), 
                      by.y = c("submission_date", "state"))
tweets_cases_vaccine <- merge(tweets_cases, vaccine, by.x = c("date", "user_state"), 
                              by.y = c("Date", "Location"), all.x = TRUE)

#write.csv(tweets_cases_vaccine, "tweets_cases_vaccine.csv", row.names=FALSE)
```








# Part 2: Sentiment Analysis

```{r}
# Load the merged data
tweets_cases_vaccine <- read.csv("tweets_cases_vaccine.csv", header = TRUE)

# Remove unnecessary words
tweets_cases_vaccine$text <- gsub("covidvaccine", "", tweets_cases_vaccine$text)
tweets_cases_vaccine$text <- gsub("covid19", "", tweets_cases_vaccine$text)
tweets_cases_vaccine$text <- gsub("covid", "", tweets_cases_vaccine$text)
tweets_cases_vaccine$text <- gsub("vaccine", "", tweets_cases_vaccine$text)

```

## Sentiment Analysis

```{r}
syuzhet <- get_sentiment(tweets_cases_vaccine$text, method="syuzhet")
tweets_sentiment <- add_column(tweets_cases_vaccine, sentiment = syuzhet, .after = "text")
colnames(tweets_sentiment) <- tolower(colnames(tweets_sentiment))
tweets_sentiment[is.na(tweets_sentiment)] <- 0
#write.csv(tweets_sentiment, "tweets_sentiment.csv", row.names=FALSE)
```

```{r}
# Compute total sentiment score and save the data (takes a lot of time!!!!)
tweets <- iconv(tweets_cases_vaccine$text)
senti <- get_nrc_sentiment(tweets)
#write.csv(senti, "nrc_sentiment.csv", row.names=FALSE)
```

## Visualize most frequent words

```{r}
wordcloud(tweets_cases_vaccine$text, min.freq = 1000, 
          colors = brewer.pal(8, "Dark2"),
          random.color = TRUE, 
          random.order = F, 
          max.words = 50, 
          scale = c(5, 0.3))
```

## Visualize sentiment scores

```{r}
# calculate total score for each sentiment
senti <- read.csv("nrc_sentiment.csv", header = T)
senti_score <- data.frame(colSums(senti[,]))
names(senti_score) <- "Score"
senti_score <- cbind("sentiment" = rownames(senti_score), senti_score)
rownames(senti_score) <- NULL
senti_score <- subset(senti_score, sentiment!="positive" & sentiment!="negative")


#plotting the sentiments with scores
ggplot(data = senti_score, aes(x = reorder(sentiment, -Score),y = Score)) + 
  geom_bar(aes(fill = Score), stat = "identity") +
  theme(legend.position = "none")+
  theme_light() +
  xlab("Sentiments") + 
  ylab("Scores") +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0)) +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
ggsave(path = "graphs", filename = "senti_bar.png", width = 6, height = 4)
```








# Part 3: Exploretory Data Analysis

```{r}
# Load datasets
tweets_sentiment <- read.csv("tweets_sentiment.csv", header = TRUE)
vaccine <- read.csv("covid_vaccine_admistrated.csv", header = TRUE)
cases <- read.csv("covid_cases.csv", header = TRUE)
cities <- read.csv("uscities.csv", header = TRUE)

tweets_sentiment$date <- as.Date(tweets_sentiment$date)
vaccine$Date <- as.Date(parse_date_time(vaccine$Date, 'mdY', tz = "UTC"))
cases$submission_date <- as.Date(parse_date_time(cases$submission_date, 'mdY', tz = "UTC"))

us_states <- map_data("state")
```

### Map state_id to state name

```{r}
state_names <- cities %>% select(state_id, state_name) %>% distinct()
tweets_sentiment <- merge(tweets_sentiment, state_names, by.x = "user_state", by.y = "state_id")
tweets_sentiment$state_name <- tolower(tweets_sentiment$state_name)
```

### Split data before/after the start of vaccination

```{r}
before_vac <- tweets_sentiment[tweets_sentiment$admin_per_100k == 0,]
after_vac <- tweets_sentiment[tweets_sentiment$admin_per_100k != 0,]
```

### 1. Distribution of twitter users by state - All time

```{r}
user_distribution <- tweets_sentiment %>% 
  group_by(state_name, user_name) %>%
  summarize(n =  n()) %>%
  summarize(count =  n()) %>%
  left_join(us_states, by = c("state_name" = "region"))

ggplot(data = user_distribution, mapping = aes(x = long, y = lat, group = group, fill = count)) +
  geom_polygon(color = "gray90", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) + 
  scale_fill_gradient2(low = "white", high = "#CB454A") +
  theme_map() + 
  labs(fill = "Count") +
  theme(legend.position = "right", plot.margin = unit(c(1, 1, 1, 1), "pt"))
ggsave(path = "graphs", filename = "user_dist.png", width = 6, height = 4)
```

### 2. Mean sentiment by state - All time

```{r}
senti_state <- tweets_sentiment %>% 
  group_by(state_name) %>% 
  summarize(mean_senti = mean(sentiment)) %>% 
  left_join(us_states, by = c("state_name" = "region"))

ggplot(data = senti_state, mapping = aes(x = long, y = lat, group = group, fill = mean_senti)) +
  geom_polygon(color = "gray90", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) + 
  scale_fill_gradient2(limit = c(-0.6, 0.8)) +
  theme_map() + 
  labs(fill = "Sentiment") +
  theme(legend.position = "right", plot.margin = unit(c(1, 1, 1, 1), "pt"))
ggsave(path = "graphs", filename = "senti_state_all.png", width = 6, height = 4)
```

### 3. Mean Sentiment by state - Before Vaccination

```{r}
senti_state_before <- before_vac %>% 
  group_by(state_name) %>% 
  summarize(mean_senti = mean(sentiment)) %>% 
  left_join(us_states, by = c("state_name" = "region"))

senti_state_bef <- ggplot(data = senti_state_before, mapping = aes(x = long, y = lat, group = group, fill = mean_senti)) +
  geom_polygon(color = "gray90", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) + 
  scale_fill_gradient2(limits=c(-0.6, 0.8)) +
  theme_map() + 
  labs(subtitle = "Before Mass Vaccination", fill = "Sentiment") +
  theme(legend.position = "bottom", plot.margin = unit(c(1, 1, 1, 1), "pt"))
senti_state_bef
ggsave(path = "graphs", filename = "senti_state_before.png", width = 6, height = 4)
```

### 5. Mean Sentiment by state - After Vaccination

```{r}
senti_state_after <- after_vac %>% 
  group_by(state_name) %>% 
  summarize(mean_senti = mean(sentiment)) %>% 
  left_join(us_states, by = c("state_name" = "region"))

senti_state_af <- ggplot(data = senti_state_after, mapping = aes(x = long, y = lat, group = group, fill = mean_senti)) +
  geom_polygon(color = "gray90", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) + 
  scale_fill_gradient2(limits=c(-0.6, 0.8)) +
  theme_map() + 
  labs(subtitle = "After Mass Vaccination", fill = "Sentiment") +
  theme(legend.position = "bottom", plot.margin = unit(c(1, 1, 1, 1), "pt"))
senti_state_af
ggsave(path = "graphs", filename = "senti_state_after.png", width = 6, height = 4)

# combine before and after plots together
ggarrange(senti_state_bef, senti_state_af, ncol = 2, nrow = 1)
ggsave(path = "graphs", filename = "senti_state_b+a.png", width = 12, height = 4)
```

### 5. Time Series for Mean Sentiment Trend

```{r}
tweets_sentiment %>% 
  group_by(date) %>% 
  summarize(mean_senti = mean(sentiment)) %>%
  ggplot(aes(x = date, y = mean_senti, color = mean_senti)) + 
  geom_point() +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Date", y = "Mean Sentiment", color = "Sentiment") +
  geom_smooth(span = 0.4, color = "dark green") +
  geom_vline(xintercept = as.Date("2020-12-25"), linetype="dashed", color = "black", size=1) +
  geom_label(aes(as.Date("2020-12-25"), 1.2), label = "Vaccination Start", show.legend = FALSE)
ggsave(path = "graphs", filename = "senti_time_all.png", width = 6, height = 4)
```

### 6. Covid Cases and Death Trend

```{r}
cases %>% group_by(submission_date) %>%
  summarise(case = mean(new_case)) %>%
  ggplot(aes(x = submission_date, y = case, color = case)) + 
  geom_point() +
  geom_smooth(span = 0.4, color = "dark green") +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Date", y = "New Cases", color = "Mean New Cases") +
  theme(legend.position = "right", plot.margin = unit(c(1, 1, 1, 1), "pt"))
ggsave(path = "graphs", filename = "case_time.png", width = 6, height = 4)

cases %>% group_by(submission_date) %>%
  summarise(case = mean(new_death)) %>%
  ggplot(aes(x = submission_date, y = case, color = case)) + 
  geom_point() +
  geom_smooth(span = 0.4, color = "dark green") +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Date", y = "New Deaths", color = "Mean New Deaths") +
  theme(legend.position = "right", plot.margin = unit(c(1, 1, 1, 1), "pt"))
ggsave(path = "graphs", filename = "death_time.png", width = 6, height = 4)
```

### 7. Vaccination rates trend

```{r}
vaccine <- vaccine %>% group_by(Location) %>%
  arrange(Location, Date) %>%
  mutate(Admin_New = rollapply(Admin_Per_100K, 2, diff , align = 'right', fill = NA))

vaccine %>% group_by(Date) %>%
  summarise(admin = mean(Admin_New)) %>%
  ggplot(aes(x = Date, y = admin, color = admin)) + 
  geom_point() +
  geom_smooth(span = 0.4, color = "dark green") +
  theme_classic() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "Date", y = "Administrated Per 100K", color = "Admin Per 100K") +
  theme(legend.position = "right", plot.margin = unit(c(1, 1, 1, 1), "pt"))
ggsave(path = "graphs", filename = "vac_time.png", width = 6, height = 4)
```

```{r}
# save admin_new as a new feature
#vaccine <- subset(vaccine, select = c(Date, Location, Admin_New))
#tweets_sentiment <- merge(tweets_sentiment, vaccine, by.x = c("date", "user_state"), 
#                              by.y = c("Date", "Location"), all.x = TRUE)
#tweets_sentiment$Admin_New[is.na(tweets_sentiment$Admin_New)] = 0
#write.csv(tweets_sentiment, "tweets_sentiment.csv", row.names = FALSE)
```








# Part 4: Modelling

```{r}
# Load final data
tweets_sentiment <- read.csv("tweets_sentiment.csv", header = TRUE)
vaccine <- read.csv("covid_vaccine_admistrated.csv", header = TRUE)
cases <- read.csv("covid_cases.csv", header = TRUE)
cities <- read.csv("uscities.csv", header = TRUE)

tweets_sentiment$date <- as.Date(tweets_sentiment$date)
vaccine$Date <- as.Date(parse_date_time(vaccine$Date, 'mdY', tz = "UTC"))
cases$submission_date <- as.Date(parse_date_time(cases$submission_date, 'mdY', tz = "UTC"))

summary(tweets_sentiment)
```

## Count observations (user + date)

```{r}
tweets_obs <- tweets_sentiment %>% group_by(user_state, date) %>% 
  summarize(sentiment = mean(sentiment))
  
cases <- subset(cases, select = c(submission_date, state, tot_cases, new_case, tot_death, 
                                  new_death))
vaccine <- subset(vaccine, select = c(Date, Location, Admin_Per_100K))

# merge and save the merged data
tweets_obs <- merge(tweets_obs, cases, by.x = c("date", "user_state"), 
                      by.y = c("submission_date", "state"))
tweets_obs <- merge(tweets_obs, vaccine, by.x = c("date", "user_state"), 
                              by.y = c("Date", "Location"), all.x = TRUE)

colnames(tweets_obs) <- tolower(colnames(tweets_obs))
tweets_obs$admin_per_100k[is.na(tweets_obs$admin_per_100k)] = 0

before_vac <- tweets_obs[tweets_obs$admin_per_100k == 0,]
after_vac <- tweets_obs[tweets_obs$admin_per_100k != 0,]

summary(tweets_obs)
```

## Covariance Matrix

```{r}
cov(after_vac %>% select("admin_per_100k", "new_case", "new_death", "sentiment", "tot_cases", "tot_death"))
```

## Basic model

```{r}
mod0 <- lm(admin_per_100k ~ sentiment, data = after_vac)
summary(mod0)
```



```{r}
mod1 <- lm(admin_per_100k ~ sentiment + user_state - 1,
           data = after_vac)
summary(mod1)
```



```{r}
mod2 <- lm(admin_per_100k ~  new_death + new_case + sentiment + user_state - 1, 
           data = after_vac)
summary(mod2)
```

```{r}
mod3 <- lm(admin_per_100k ~  new_death*new_case + sentiment + user_state - 1, 
           data = after_vac)
summary(mod3)
```


```{r}
library(stargazer)

rob_se <- list(sqrt(diag(vcovHC(mod0, type = "HC1"))),
               sqrt(diag(vcovHC(mod1, type = "HC1"))),
               sqrt(diag(vcovHC(mod2, type = "HC1"))),
               sqrt(diag(vcovHC(mod3, type = "HC1"))))

stargazer(mod0, mod1, mod2, mod3,
          digits = 3,
          header = FALSE,
          type = "latex", 
          se = rob_se,
          title = "Linear Regression Models",
          model.numbers = FALSE,
          column.labels = c("(1)", "(2)", "(3)", "(4)"))

```
